\section{Mathematical Logic}

So we start our discussion of logic with an idea.
Now ideas can be seen to have two distinct values of validity: true and false.
We refine these ideas into a statement, which is defined as follows:

\begin{defn}
	\index{Statement}
    \index{Proposition|see{Statement}}
    \label{defn:statement}
	A \underline{\emph{statement}} is a phrase which is either true or false. (A statement can be represented by a letter for simplicity and compactness.)
\end{defn}

So a statement encompasses an entire idea and can be evaluated to determine its validity

\begin{ex}
    For instance we have the following statements:
	\begin{center}
    P="The sky is blue".\\
	Q="The sky is brown".
    \end{center}
    We see the statement P is \emph{True} while the statement Q is \emph{False}.
\end{ex}

So what if our statement is simply the opposite of another statement, how can we easily represent it?
We have invented notation for this exact purpose, it is called the negation and is defined like so:

\begin{defn}
	\index{Negation $\neg$}
	\nomenclature{$ \land $}{Logical And. True if all elements are true and false otherwise.}
	The \underline{\emph{Negation}} ($\neg$) of a statement is true when the statement is false and false if the statement is true.
	(The statement $\neg P$ is written in other words as "Not P".)
\end{defn}

\begin{ex}
    Using the examples above:
    \begin{center}
	$\neg P=$"The sky is not blue".\\
	$\neg Q=$"The sky is not brown".
    \end{center}
    Here $\neg P$ is \emph{False} and $\neg Q$ is \emph{True}.
\end{ex}

\subsection{And and Or}
So far we have only been able to create very simple statements, but what if we want to combine multiple statements into a larger statement?
We can combine them using connectives \emph{and} and \emph{or}.
Since in language these two words are defined very vaguely, we will give them exact definitions so they may be used precisely.

\begin{defn}
	\index{And, Logical $\land$}
	\nomenclature{$ \lor $}{Logical Or. True if any elements is true and false otherwise.}
	A \underline{\emph{conjunction operator}}, or "And" ($\land$), connects logical statements which is defined to be true if all statements are true and false if any are false.
\end{defn}

To help better understand how these operators work we use a table which places all possible inputs with their corresponding outputs in a table.
This table is then called a truth table where the two possibilities are true(T) and false(F) for every field.
The truth table for the conjunction operator is as follows:

\begin{center}
	\begin{tabular}{cc|c}
		$P$ & $Q$ & $P \land Q$ \\
		\hline
		T & T & T \\
		T & F & F \\
		F & T & F \\
		F & F & F \\
	\end{tabular}
\end{center}

So what happens if we take the negation of the statement $P \land Q$?
Well, the statement says, "Not both P and Q", which can be interpreted as "neither P nor Q", which again can be rewritten as "Not P or Not Q".
Using this idea or an \emph{or} connective we can define the disjunction operator as follows:

\begin{defn}
	\index{Or, Logical $\lor$}
	A \underline{\emph{disjunction operator}}, or "Or" ($\lor$), connects logical statements and is true if any statements are true and false if all are false.
\end{defn}

So lets look at the truth table for the disjunction operator with the conjunction operator.

\begin{center}
	\begin{tabular}{cc|ccc}
		$P$ & $Q$ & $P \land Q$ & $P \lor Q$ & $\neg P \lor \neg Q$ \\
		\hline
		T & T & T & T & F \\
		T & F & F & T & T \\
		F & T & F & T & T \\
		F & F & F & F & T \\
	\end{tabular}
\end{center}

From this we can see that $\neg (P \land Q)$ is equivalent to $\neg P \lor \neg Q$.
We can also look at the negation of the statement $P \lor Q$ by looking at it as the sentence, "Not either P or Q", and the only time that will work is when both P and Q are false, so the sentence becomes, "Not P and Not Q".
This can be rewritten as $\neg P \land \neg Q$.
This brings us to our first theorem, originally stated as laws by De Morgan.

\begin{thm}
	\index{De Morgan's Laws (Logic)}
	\textbf{(De Morgan's Laws)} \\
	\label{Thm:DeMorganLogic}
	\begin{equation}
		\nonumber
		\neg ( P \land Q ) \iff \neg P \lor \neg Q
	\end{equation}
	\begin{equation}
		\nonumber
		\neg ( P \lor Q ) \iff \neg P \land \neg Q
	\end{equation}
\end{thm}
Here the symbol $\iff$ means logically equivalent, we will generalize its meaning to "if and only if" operator later in the chapter.

\begin{proof}
	To prove theorem \vref{Thm:DeMorganLogic} we can construct the truth tables for the statements involved relatively simply.
	\begin{table*}[ht]
	\begin{center}
	\begin{tabular}{cc|cc|cccccc}
			$P$ & $Q$ & $\neg P$ & $\neg Q$ & $P \land Q$ & $\neg (P \land Q)$ & $P \lor Q$ & $\neg (P \lor Q)$ & $\neg P \lor \neg Q$ & $\neg P \land \neg Q$ \\
			\hline
			T & T & F & F & T & F & T & F & F & F \\
			T & F & F & T & F & T & T & F & T & F \\
			F & T & T & F & F & T & T & F & T & F \\
			F & F & T & T & F & T & F & T & T & T \\
	\end{tabular}
	\end{center}
	\end{table*}
	Since there are only four possible inputs for these operators, all possibilities have been shown and the theorem is proved.
\end{proof}

\begin{thm}
	\textbf{Properties of Logical Operators}\\
	\label{thm:proplogical}
	\textbf{Distributivity}: $(A \land B) \lor C \iff (A \lor C) \land (B \lor C)$ and $(A \lor B) \land C \iff (A \land C) \lor (B \land C)$
	
\end{thm}

\subsection{Implication}
We are ultimately concerned with using logic to show how two idea are connected.
Usually this connectedness is expressed in the form of a conditionals of the form, "If \emph{something} then \emph{something else}."
In terms of logic when a true statement causes another to also be true an implication is formed.
We can now define the implication as follows:

\begin{defn}
	\index{Implication, $ \Rightarrow $}
	An \emph{\underline{implication}} ($A \Rightarrow B$) between means that if A is true, so must B.
\end{defn}
With this definition, if A is false then B is unrestricted and can be anything.
So if A is false, the statement will always be true since a false statement implies anything.
We also note that if A is true but B is false, that would suggest that A doesn't imply B, so the statement is false.
Using this we can construct a truth table for an implication.

\begin{center}
	\begin{tabular}{cc|c}
		$P$ & $Q$ & $P \Rightarrow Q$ \\
		\hline
		T & T & T \\
		T & F & F \\
		F & T & T \\
		F & F & T \\
	\end{tabular}
\end{center}

So if $A \Rightarrow B$ is true, then a false B would imply a false A.
This brings us to the definition of the contrapositive.

\begin{defn}
	\index{Contrapositive}
	Given $A \Rightarrow B$, $\neg B \Rightarrow \neg A$ is called the contrapositive of $A \Rightarrow B$.
\end{defn}

The Contrapositive and the implication related to it are equivalent.
